---
title: "Makeover Monday - Deloitte Christmas Survey 2019"
author: "Billy Caughey"
date: "12/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

## Introduction

This weeks Makeover Monday is from the [Deloitte Christmas Survey 2019]("https://www2.deloitte.com/uk/en/pages/consumer-business/articles/deloitte-christmas-survey-2019.html"). The visual is a representation of the spend in 2018 compared to the spend in 2019. This representation is done in a table with arrows showing the precent change. 

This is a nice representation but I am going to try to present the information in a different way. As a statistician, histograms are great methods to use and incorporate. This is because we can can a lot of information in a preceise way.

## Data Import

Again, the folks at [data.world]("https://data.world/makeovermonday/2019w52") have posted the data will an R code to pull the data down to R.

```{r data import, results = "hide"}

library(httr)
library(readxl)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(scales)
GET("https://query.data.world/s/yuycdlchnfcjcp35gc4zxzqsfhxduf", 
    write_disk(tf <- tempfile(fileext = ".xlsx")))
df <- read_excel(tf)

```

The first plot I will make is a barplot showing the amount spent (in pounds) in each category by year. I will use a facet to show the UK and Europe against each other.

```{r bar1, fig.width = 10}


bar1 <- df %>%
    ggplot(aes(x = Category, y = `Amount in £`, fill = as.character(Year))) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(fill = "Year") +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 14),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 12),
          strip.text = element_text(size = 20),
          legend.position = "top") +
    facet_wrap(~ Region, nrow = 1) 

bar1

```

The second plot, I will use another barplot. This time, I will do some footwork to compute the percent change from year to year in each category for the UK and Europe. I plot using the same method as the previous plot without use of a facet. This will let me make a percent change for each category within each region.

```{r bar2, fig.width = 10}

bar2 <- df %>%
    group_by(Region, Category) %>%
    arrange(Year, .by_group = TRUE) %>%
    mutate(`Pct Change` = (`Amount in £`/lag(`Amount in £`)) - 1) %>%
    select(Region, Category, `Pct Change`) %>%
    filter(is.na(`Pct Change`) == FALSE) %>%
    ggplot(aes(x = Category, y = `Pct Change`, fill = Region)) + 
    geom_bar(stat = "identity", position = "dodge") +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 14),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 12),
          legend.position = "top") +
    scale_y_continuous(labels = percent) 

bar2

```

Finally, I make use of the `ggarrange` function. This function allows the user to combine plots made from the `ggplot` library. A similar method can be found in base R using the `par(mfrow())` method. I should note I change the colors for the year to year comparison. It did not make sense to use the same colors in both charts. In fact, I would argue it makes it difficult to the reader. One additional thing, I use the `theme_bw` option. In my personal opinion, plots look better in `theme_bw`.

```{r bar1 + bar2, fig.width = 10, fig.height = 8}

bar1 <- df %>%
    ggplot(aes(x = Category, y = `Amount in £`, fill = as.character(Year))) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(fill = "Year") +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 14),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 12),
          strip.text = element_text(size = 20),
          legend.position = "top") +
    scale_fill_manual(values=c("#999999", "#E69F00")) +
    facet_wrap(~ Region, nrow = 1) 

bar2 <- df %>%
    group_by(Region, Category) %>%
    arrange(Year, .by_group = TRUE) %>%
    mutate(`Pct Change` = (`Amount in £`/lag(`Amount in £`)) - 1) %>%
    select(Region, Category, `Pct Change`) %>%
    filter(is.na(`Pct Change`) == FALSE) %>%
    ggplot(aes(x = Category, y = `Pct Change`, fill = Region)) + 
    geom_bar(stat = "identity", position = "dodge") +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 14),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 12),
          legend.position = "top") +
    scale_y_continuous(labels = percent) 

bar3 <- ggarrange(bar1, bar2, nrow = 2, align = "hv")

bar3

```

This is my submission for this week.

## Data Analysis

Unforunately, the data set is rather small. This does not allow for a statistical analysis at this point. I recommend the reader goes to Deliotte's survey and read their high level report.

