---
title: "Makeover Monday, 2020w9"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE, message=F, comment=F, warning=F}

##### Libraries #####

library(flexdashboard)
library(httr)
library(readxl)
library(tidyverse)
library(kableExtra)

##### Data Import #####

school_age = function(x){
    
    if(x == "Kindergarten"){return(6)}
    else if(x == "First grade"){return(7)}
    else if(x == "Second grade"){return(8)}
    else if(x == "Third grade"){return(9)}
    else if(x == "Fourth grade"){return(10)}
    else if(x == "Fifth grade"){return(11)}
    else if(x == "Sixth grade"){return(12)}
    else if(x == "Seventh grade"){return(13)}
    else if(x == "Eighth grade"){return(14)}
    else if(x == "Ninth grade"){return(15)}
    else if(x == "10th grade"){return(16)}
    else if(x == "11th grade"){return(17)}
    else {return(18)}

}

invisible(capture.output(GET("https://query.data.world/s/hf5w6pnh5rxzbg5uz7qzc5a7p2cq5k", 
    write_disk(tf <- tempfile(fileext = ".xlsx")))))
df <- read_excel(tf) %>%
    mutate(diff = `Hours Needed` - `Hours Averaged`,
           age = sapply(`Grade`, school_age))


mod1 <- lm(diff ~ age, data = df)
tab1 <- round(summary(mod1)$coefficients, 3) %>%
    kable(caption = "Table 1. Linear Regression Results") %>%
    kable_styling()


```

Column {data-width=500}
-----------------------------------------------------------------------

### Bar Graph of Hours Slept vs Hours Needed

```{r}

df %>%
    select(`Grade`, `Hours Needed`, `Hours Averaged`) %>%
    mutate(`Grade` = factor(x = `Grade`,
                            levels = c("12th grade","11th grade","10th grade",
                                       "Ninth grade","Eighth grade","Seventh grade",
                                       "Sixth grade","Fifth grade","Fourth grade",
                                       "Third grade","Second grade","First grade",
                                       "Kindergarten"))) %>%
    gather(key = `Hours`,
           value = `Sleep`,
           `Hours Needed`, `Hours Averaged`) %>%
    ggplot(aes(x = `Grade`, y = `Sleep`, fill = `Hours`)) +
    geom_bar(stat = 'identity', position = 'dodge') +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.text.y = element_text(size = 14),
          legend.position = "bottom") +
    coord_flip() 

```



Column {data-width=500}
-----------------------------------------------------------------------

### Scatterplot of Age vs Slept Hours Missed

```{r}

df %>%
    select(`Grade`, `diff`, `age`) %>%
    ggplot(aes(x = `age`, y = `diff`)) +
    xlab("Student Age") +
    ylab("Hours Missed") +
    geom_point(size = 2) +
    stat_smooth(method = lm) +
    theme_bw() +
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))


```

### Results 




`r tab1`

Over time, the difference between the amount needed and the amount obtained is increasing. A students year in school was translated to a student's age. On average, a student at any time is losing `r round(summary(mod1)$coefficients, 3)[1,1]` hours. Incrementally, as the students age increases, the difference increases by `r round(summary(mod1)$coefficients, 3)[2,1]` hours ( p = 0.001). 

These results are evidence to consider the factors pointed out savvysleeper's [article](https://savvysleeper.org/costing-kids-sleep/) seriously. 

Data source: https://data.world/makeovermonday/2020w9

