---
title: "Makeover Monday 2021, Week 2"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}

##### Libraries #####

library(flexdashboard)
library(tidyverse) 
library(Amelia)

##### Pre-Processing #####

file_url <- "https://query.data.world/s/2bnbe3zaxn2w2w72xisuwcenvjx5r5"
df <- read.csv(file_url, 
               header=TRUE, 
               stringsAsFactors=FALSE) 

## Review structure ##

str(df)

## Put it together ##

df1 <- df %>%
    rename(region = UNICEF.Region,
           new_hiv_infection_per_1000 = Estimated.incidence.rate.of.new.HIV.infection.per.1.000.uninfected.population,
           gender = Sex,
           year = Year,
           aids_related_deaths_per_100000 = Estimated.rate.of.annual.AIDS.related.deaths..per.100.000.population) %>%
    select(region, gender, year, 
           new_hiv_infection_per_1000, aids_related_deaths_per_100000) %>%
    group_by(region, gender, year) %>%
    summarize(new_hiv_infection_per_1000_avg = mean(new_hiv_infection_per_1000, na.rm = T),
              aids_related_deaths_per_100000_avg = mean(aids_related_deaths_per_100000, na.rm = T)) %>%
    mutate(region = ifelse(region == "Eastern and Southern Africa",
                           "Eastern and \nSouthern Africa",
                           "West and \nCentral Africa"))


```

This week's makeover is about HIV infections in Africa. Specifically, we consider new HIV infections and AIDs related deaths. A few key points from this information:

- Young ladies, specifically in Eastern and Southern Africa, have the highest new infection rates from 1990-2020. These infection rates, although decreasing, are still higher than any other new infection rate.
- The lag in AIDs related deaths coincide with the high infection rates in the 1990's. 
- Eastern and Southern Africa have higher aides related deaths per 100,000 from 1990 - 2020. These rates peaked in the mid 2000's, but are still higher than West and Central Africa.
- Within the Eastern and Southern African regions, young women have a higher death rate than young men do. 

Based on these figures, the young women of Africa need help and support. 



Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### New Infections

```{r}

df1 %>%
    ggplot(aes(x = year, y = new_hiv_infection_per_1000_avg)) +
    geom_line(size = 2) +
    theme_bw() +
    labs(y = "New Infections Per 1,000 (Avg)",
         title = "New HIV Infections Per 1,000 (Avg)\n1990 - 2020") +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 45, vjust = 0.5),
          axis.title.y = element_text(size = 16),
          axis.text.y = element_text(size = 12, angle = 45, vjust = 0.5),
          strip.text = element_text(size = 16),
          plot.title = element_text(size = 18)) +
    facet_grid(region ~ gender)

```

### Related Deaths

```{r}

df1 %>%
    ggplot(aes(x = year, y = aids_related_deaths_per_100000_avg)) +
    geom_line(size = 2) +
    theme_bw() +
    labs(y = "Deaths per 100,000 (Avg)",
         title = "Aids related deaths per 100,000 (Avg)\n1990 - 2020") +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12, angle = 45, vjust = 0.5),
          axis.title.y = element_text(size = 16),
          axis.text.y = element_text(size = 12, angle = 45, vjust = 0.5),
          strip.text = element_text(size = 16),
          plot.title = element_text(size = 18)) +
    facet_grid(region ~ gender)

```

