---
title: "Monday Makeover 2020w32"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}

##### Library ##### 

library(tidyverse)
library(httr)
library(readxl)
library(lubridate)
library(ggpubr)
library(scales)

##### Bring in the data #####

url <- "https://query.data.world/s/otvxdqu7wcrxpzwz6vikc3vxxc2lld"
invisible(capture.output(GET(url, write_disk(tf <- tempfile(fileext = ".xlsx")))))
df <- read_excel(tf)

##### Preprocessing #####

df_bene <- df %>%
  mutate(`Benefit of remote work` = factor(`Benefit of remote work`,
                                           levels = c("Other",
                                                      "Ability to work from home",
                                                      "Ability to spend time with family",
                                                      "Not having to commute",
                                                      "Flexibility to work from any location",
                                                      "Ability to have a flexible schedule")))

## Add the struggles 

struggle <- c("Difficulties with collaboration and communication",
              "Loneliness",
              "Not being able to unplug",
              "Distractions at home",
              "Being in a different timezone than teammates",
              "Staying motivated",
              "Taking vacation time",
              "Finding reliable wifi",
              "Other")

percent <- c(20, 20, 18, 12, 10, 7, 5, 3, 5)

df1 <- tibble(`Percentage` = percent,
              `Struggles of remote work` = struggle) %>%
  mutate(`Struggles of remote work` = factor(`Struggles of remote work`,
                                             levels = c("Other",
                                                        "Finding reliable wifi",
                                                        "Taking vacation time",
                                                        "Staying motivated",
                                                        "Being in a different timezone than teammates",
                                                        "Distractions at home",
                                                        "Not being able to unplug",
                                                        "Loneliness",
                                                        "Difficulties with collaboration and communication")))

##### Visual #####

# benefits

bene_viz <- ggplot(df_bene, 
                   aes(x = `Benefit of remote work`, 
                       y = `Percentage`)) +
  geom_bar(stat = "identity") +
  ggtitle("Benefits of Working at Home") +
  theme_bw() +
  theme(legend.key = element_blank(),
           axis.title.x = element_blank(),
           axis.text.x = element_text(size = 12),
           axis.title.y = element_blank(),
           axis.text.y = element_text(size = 12)) +
  coord_flip() 

# Struggles

strg_viz <- ggplot(df1, 
                   aes(x = `Struggles of remote work`, 
                       y = `Percentage`)) +
  geom_bar(stat = "identity") +
  ggtitle("Struggles of Working at Home") +
  ylim(c(0,32)) +
  theme_bw() +
  theme(legend.key = element_blank(),
           axis.title.x = element_blank(),
           axis.text.x = element_text(size = 12),
           axis.title.y = element_blank(),
           axis.text.y = element_text(size = 12)) +
  coord_flip() 


```


### Benefits vs Struggles

The benefits of working from home seem to have some powerful benefits. For instance, the ability to have a flexible schedule seems to be at the foremost of everyone's mind. What is interesting is the top three benefits and the top three struggles seem to be related. For starters, having a flexible schedule means the employee has the ability to jump on at any time. So, if for example, the employee is taking a trip, they can jump on at the end of their trip. This also lends itself to being able to work from anywhere without a commute! Alternatively, working from home can tax the ability to communicate with friends and colleagues at work. This definitely can lead to loneliness. Loneliness leads the individual to do anything at their disposal... which just can happen to be work. Thus, the three top benefits and struggles seem to balance each other out with percent of responses.

```{r, echo = F, fig.width = 14}

ggarrange(bene_viz, 
          strg_viz, 
          ncol = 1, 
          align = "hv")

```

