---
title: "Makeover Monday 2020w27"
author: "Billy Caughey"
date: "July 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      comment = FALSE)
```

## Background

This is the only section I will use first person as I am speaking directly to the reader. As an R programming enthusiast, I find my participation in Makeover Monday more of a coding exercise rather than a dashboarding exercise. In the spirit of Makeover Monday, I do what I can to remain consistent with visual analytics and work with the `ggplot2` library. In my experience, data visualization can do a large majority of the analytic lift. I would argue 75-85% of the lift. With that said, the majority of my submissions and explorations are visual in nature. With that said, when the opportunity to exercise statistics, machine learning, artificial intelligence, etc. arises in these exercises, I will take them!

As always, please feel free to leave a comment. I would enjoy the constructive criticism and how I can improve!

## Introduction

This week's data is around [mental health in the United Kingdom](https://www.bbc.com/news/health-41125009). Men and Women were surveyed to about their mental health status. The survey showed women are more likely to have a mental health condition across all age groups. The goal of this week is to see if there is an improvement which can be made on the visual.

## Critiques of the visual

This weeks visual shows the proportion of people who have mental disorder symptoms in the last week.

```{r week 23 viz, fig.align = 'center', out.width = "75%", echo = FALSE, fig.cap = "Fig. 1: Viz of the week", cache = TRUE}

library(knitr)
library(png)

img1_path <- "makeovermonday2020w27.png"

include_graphics(img1_path)

```

### The Good...

The visual does a great job at showing the porportion of men and women who have symptoms. The proportions are easily read and understood. A smaller note should be the colors used. They do not clash and are very easy on the eyes.

### What could be improved...

A second chart or an additional bar showing the difference between groups can be added to improve the chart. This difference can be used as a measure to understand the magnitude of difference in each age group. If a new chart is used, neutral colors should be used. If on the same chart, a complimentary color should be used.

### There's a statistics option here...

On the statistics front, there is a hypothesis to be tested. This hypothesis is there is not an association between gender and age category in proprotions reporting mental health symptoms. For this test, an $\alpha$ of 0.05 will determine statistical significance.

## New visual

The first step of the new visual is bringing in the data. Like all other Makeover Monday's, the data is stored on [data.world](https://data.world/makeovermonday/2020w27-comparing-common-mental-disorder-by-sex). The code to import the data is below.

```{r data import}

library(tidyverse)
library(httr)
library(readxl)

url <- "https://query.data.world/s/6c67766li2iawiudv3nl2xt32xkihv"
invisible(capture.output(GET(url, 
                             write_disk(tf <- tempfile(fileext = ".xlsx")))))

df <- read_excel(tf) %>%
  gather(key = age_category, value = proportion,
         `16-24`, `25-34`, `35-44`, `45-54`, `55-64`, `65-74`, `75+`, `All`)

str(df)

```