---
title: "Makeover Monday 2020w2 Exploration"
author: "Billy Caughey"
date: "June 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
<<<<<<< HEAD
knitr::opts_chunk$set(echo = TRUE, message = F, comment = F, warning = F)
=======
knitr::opts_chunk$set(echo = TRUE, message = F, comment = F, warnings = F)
>>>>>>> eaaf344af0a84bbb3a9d2d9e7e0fbce41815b649
```

## Introduction

This is my submission for week 2 of the Makeover Monday social project. This week's topic revolves around the use of pesticides in the United States. More specifically, the purpose is to compare the pesticides used in the United States which considered banned in other countries. 

The "viz" which will be critiqued and compared is below in figure 1.

```{r mm viz, fig.align = 'center', out.width = "75%", echo = FALSE, fig.cap = "Fig. 1: Makeover Monday W2 Viz", cache = TRUE}

library(knitr)
library(png)

img1_path <- "makeover_monday_wee2_challenge.png"

include_graphics(img1_path)

```

## What I like about the viz

The visual is straight forward. There aren't a lot of bells and whistles to sift through. The x- and y-axis are labeled clearly. Additionally, it is very clearly seen which countries are included in this comparison.

## What I think I can improve 

Although this visual is rather straight forward, I would like to know more about the cross over. There are 85 pesticides among the 3 countries which have been banned. What is the distribution of these pesticides among these countries? I would love to see a Venn diagram in this situation. It might look ridiculous, but it would be informative. 

With that said, I think I can distill enough information to be able to support the build of a Venn diagram. If I can do this, then I will be able to build a neat Venn diagram.

## Bringing in the data 

Okay, let's bring in the data and see what we have going on.

```{r data import}

library(httr)
library(readxl)
url <- "https://query.data.world/s/h2y64h7jm7wgvux5hwyumkbnrhvuop"
invisible(capture.output(GET(url, write_disk(tf <- tempfile(fileext = ".xlsx")))))
df <- read_excel(tf)
print(df)

```

It looks like there has been a fast one pulled. In the visual, the number of pesticides are being compared. This data is comparing the poundage being used. It shouldn't matter, but it is a touch frustrating. This is something most data folks have to get used to in practice. Sometimes bad or miscommunicated data comes through.

## Trying to build a Venn diagram

The Venn diagram I will try to build is something like figure 2.

```{r venn diagram, fig.align = 'center', out.width = "75%", echo = FALSE, fig.cap = "Fig. 1: Venn Diagram", cache = TRUE}

library(knitr)
library(png)

img1_path <- "venn diagram.png"

include_graphics(img1_path)

```

A decision has to be made on the banned substance that is 'unknown' or 'banned in all 3'. Since I can't confidently say it's unknown or all 3, I'm going to exclude this information at this point. This means the data will look like the following.

```{r data update}

library(tidyverse)

denom <- df$`Lbs. Pesticides Used in USA Agriculture`[1] - df$`Lbs. Pesticides Used in USA Agriculture`[8]

df2 <- df %>%
  filter(!(Category %in% c("Total","Banned, Not Approved or Unknown in All 3"))) %>%
  mutate(new_perc = round((`Lbs. Pesticides Used in USA Agriculture` / denom) * 100, 1))

print(df2)

```

<<<<<<< HEAD
With the new data in, we need to determine which pesticides are coming between groups. The best way to do that is to use the `VennDiagramm` library. In this library, we will use the `venn.diagram` function. The first venn I will create is the number of pesticides all three groups are banning or phasing out.

```{r venn diagram, echo = F}

library(VennDiagram)

venn.diagram(x = list(dat_eu$Pesticide, dat_ch$Pesticide, dat_br$Pesticide),
             category.names = c("EU","CH","BR"),
             filename  = "2020w2_venn_diagram.png",
             output = TRUE,
             na = "remove")

```

```{r plot venn diagram}

library(knitr)
library(png)

img2_path <- "2020w2_venn_diagram.png"

include_graphics(img2_path)

```
=======
With this information, I can write out what is known to build out the equations I will use. 

$$
EU = A + x + y + w = 30.2 \\
Brazil = B + y + z + w = 2.4 \\
China = C + x + z + w = 3.8 \\
1+ = A + B + C + x + y + z + w = 30.7 \\
2+ = x + y + z + w = 4.3 \\\
3 = w = 1.4
$$

Since `w` is know, another equation will need to be writtens up. Unfortunately, any new function will be a linear combination of the previous data. This will require a more indepth review of the data.

In reviewing the actual [article](https://ehjournal.biomedcentral.com/articles/10.1186/s12940-019-0488-0#Tab1), all the source files are cited. The source file needed for this analysis will be [additional file 5](https://static-content.springer.com/esm/art%3A10.1186%2Fs12940-019-0488-0/MediaObjects/12940_2019_488_MOESM5_ESM.xlsx). Specifically, we will use sheets `S131`, `s132`, and `s133` for this visual.  
>>>>>>> eaaf344af0a84bbb3a9d2d9e7e0fbce41815b649

```{r }






