---
title: "Makeover Monday, 2020w8"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}

##### Libraries #####

library(flexdashboard)
library(httr)
library(readxl)
library(tidyverse)
library(scales)
library(ggpubr)
library(kableExtra)

##### Pulling in the data #####

invisible(capture.output(GET("https://query.data.world/s/btmtcqnqumda6yhsxzqwtzbyb5uar7", 
    write_disk(tf <- tempfile(fileext = ".xlsx")))))
df <- read_excel(tf, sheet = 2)

##### Explorative Data Analytsis #####

df1 <- df %>%
    mutate(completion = ifelse(`Client situation at end of support` == "Unknown",
                               "Dropped Out","Completed")) %>%
    group_by(Year, completion) %>%
    summarise(grp_obs = sum(`Size (no. of clients)`)) %>%
    mutate(pct_complete = grp_obs / sum(grp_obs)) %>%
    ungroup()

comp1415 <- df1 %>%
    filter(Year == names(table(df1$Year))[1]) %>%
    ggplot(aes(x = completion, y = pct_complete)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2014-15")  +
    coord_flip()

comp1516 <- df1 %>%
    filter(Year == names(table(df1$Year))[2]) %>%
    ggplot(aes(x = completion, y = pct_complete)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2015-16")  +
    coord_flip()

comp1617 <- df1 %>%
    filter(Year == names(table(df1$Year))[3]) %>%
    ggplot(aes(x = completion, y = pct_complete)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2016-17")  +
    coord_flip()

comp1718 <- df1 %>%
    filter(Year == names(table(df1$Year))[4]) %>%
    ggplot(aes(x = completion, y = pct_complete)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2017-18")  +
    coord_flip()


##### homeless situation #####

df2 <- df %>%
    mutate(`Housing situation at end of support` = ifelse(
        `Client situation at end of support` %in% c("Couch surfer","Rough sleeper","Short term temporary accommodation"),"Homeless","At Risk")) %>%
    select(Year, `Housing situation at start of support`, `Housing situation at end of support`, `Size (no. of clients)`) %>%
    mutate(situation = paste0(`Housing situation at start of support`," - \n", `Housing situation at end of support`)) %>%
    group_by(Year, situation) %>%
    summarize(Size = sum(`Size (no. of clients)`)) %>%
    mutate(pct_situation = Size / sum(Size))

home1415 <- df2 %>%
    filter(Year == names(table(df1$Year))[1]) %>%
    ggplot(aes(x = situation, y = pct_situation)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2014-15")  +
    coord_flip()

home1516 <- df2 %>%
    filter(Year == names(table(df1$Year))[2]) %>%
    ggplot(aes(x = situation, y = pct_situation)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2015-16")   +
    coord_flip()

home1617 <- df2 %>%
    filter(Year == names(table(df1$Year))[3]) %>%
    ggplot(aes(x = situation, y = pct_situation)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2016-17")   +
    coord_flip()

home1718 <- df2 %>%
    filter(Year == names(table(df1$Year))[3]) %>%
    ggplot(aes(x = situation, y = pct_situation)) +
    geom_col() +
    theme_bw() +
    theme(axis.text.x = element_text(size = 8),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8),
          axis.title.y = element_blank()) +
    scale_y_continuous(labels = percent) +
    ggtitle("2017-18")   +
    coord_flip()



```


Column {data-width=500}
-----------------------------------------------------------------------

### Drop Out Rate

```{r}

ggarrange(comp1415, comp1516, comp1617, comp1718,
          ncol = 2, nrow = 2)

```

### Did the intervention work?

```{r}

ggarrange(home1415, home1516, home1617, home1718,
          ncol = 2, nrow = 2)


```





Column {data-width=500}
-----------------------------------------------------------------------

### Summary

This week's Makeover is from the Australian Specialist Homelessness Services. The data will be collected from [data.world]("https://data.world/makeovermonday/2020w8"). There are two main questions I want to answer. First, I want to know the dropout rate from year to year. The second question I want to answer is did the intervention work? 

A client will have dropped out if `Client situation at end of support == "Unknown"`.

Over the study duration, the drop out rate remained less than 8% from year to year. This is an incredible dropout rate. This also speaks to the researchers doing a fantastic job keeping connected with clients.

Now, we address if the intervention worked. At first glance, the answer may be no; however, this is not accurate. As the 'Homeless-Homeless' category begins to shirt in size, the 'Homeless-At risk' category begins to increase. So, yes, over time this program seems to be working.