---
title: "UK visits abroad: All visits Thousands - SA"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

This is my dashboard for this week's Makeover Monday. The purpose of this dash is to present additional views of UK visits abroad. The original data presented can be found [here](https://www.ons.gov.uk/peoplepopulationandcommunity/leisureandtourism/timeseries/gmax/ott). 

The trend line for total visits (red line) shows a steady increase of visits over time. Interestingly enough is the percent change over time is actually decreasing. This is best seen on the 'By Year' tab. This could suggest the number of visits are plateauing. 

```{r setup, include=FALSE, message = F, waring = F, comment = F}

##### Libraries #####

library(flexdashboard)
library(tidyverse)
library(httr)
library(readxl)
library(lubridate)
library(ggpubr)
library(scales)

##### Bring in data #####

url <- "https://query.data.world/s/igxj3pmxgtrtfum2lorsoedczkkcao"
invisible(capture.output(GET(url, write_disk(tf <- tempfile(fileext = ".xls")))))
df <- read_excel(tf) %>%
    filter(grepl(pattern = "[0-9][0-9][0-9][0-9]\\s[A-Z][A-Z][A-Z]", 
                 x = Title)) %>%
    rowwise() %>%
    mutate(year = paste0(str_split(string = Title, pattern = " ")[[1]][1]," Jan 01"),
           month = paste0(Title, " 01"),
           qtr = ifelse(grepl(pattern = "JAN|FEB|MAR", x = Title), 
                        paste0(str_split(string = Title, pattern = " ")[[1]][1]," Mar 01"),
                 ifelse(grepl(pattern = "APR|MAY|JUN", x = Title), 
                        paste0(str_split(string = Title, pattern = " ")[[1]][1]," Jun 01"),
                 ifelse(grepl(pattern = "JUL|AUG|SEP", x = Title), 
                        paste0(str_split(string = Title, pattern = " ")[[1]][1]," Sep 01"),
                        paste0(str_split(string = Title, pattern = " ")[[1]][1]," Dec 01"))))) %>%
    ungroup() %>%
    rename(uk_visits = `UK visits abroad: All visits Thousands - SA`) %>%
    mutate(year = as.Date(year, format = "%Y %b %d"),
           month = as.Date(month, format = "%Y %b %d"),
           qtr = as.Date(qtr, format = "%Y %b %d"),
           uk_visits = as.numeric(uk_visits)) %>%
    select(year, month, qtr, uk_visits) %>%
    filter(lubridate::year(year) != 2020)



```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### By Year

```{r}

by_year <- df %>%
    group_by(year) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    ggplot(aes(x = year, y = visits)) +
    geom_point(color = "red", size = 2) +
    geom_line(color = "red", size = 1) +
    geom_smooth(se = FALSE, method = "lm", color = "black") + 
    labs(y = "Visits") +
    scale_y_continuous(labels = comma) +
    theme_classic() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))
          

by_year_pct_chg <- df %>%
    group_by(year) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    arrange(year) %>%
    mutate(pct_change = (visits / lag(visits) - 1) * 100,
           pct_change = ifelse(is.na(pct_change), 0, pct_change)) %>%
    ggplot(aes(x = year, y = pct_change)) +
    geom_point(color = "blue", size = 2) +
    geom_line(color = "blue", size = 1) + 
    geom_smooth(se = FALSE, method = "lm", color = "black") +
    labs(y = "Percent Change",
         x = "Year") +
    theme_classic() + 
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))

ggarrange(by_year, 
          by_year_pct_chg, 
          ncol = 1, 
          align = "hv")

```


### By Quarter

```{r}


by_qtr <- df %>%
    group_by(qtr) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    arrange(qtr) %>%
    ggplot(aes(x = qtr, y = visits)) +
    geom_point(color = "red", size = 2) +
    geom_line(color = "red", size = 1) +
    geom_smooth(se = FALSE, method = "lm", color = "black") + 
    labs(y = "Visits") +
    scale_y_continuous(labels = comma) +
    theme_classic() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))


by_qtr_pct_chg <- df %>%
    group_by(qtr) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    arrange(qtr) %>%
    mutate(pct_change = (visits / lag(visits) - 1) * 100,
           pct_change = ifelse(is.na(pct_change), 0, pct_change)) %>%
    ggplot(aes(x = qtr, y = pct_change)) +
    geom_point(color = "blue", size = 2) +
    geom_line(color = "blue", size = 1) + 
    geom_smooth(se = FALSE, method = "lm", color = "black") +
    labs(y = "Percent Change",
         x = "Quarter") +
    theme_classic() + 
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))

ggarrange(by_qtr, 
          by_qtr_pct_chg, 
          ncol = 1, 
          align = "hv")

```

### By Month

```{r}

by_mth <- df %>%
    group_by(month) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    arrange(month) %>%
    ggplot(aes(x = month, y = visits)) +
    geom_point(color = "red", size = 2) +
    geom_line(color = "red", size = 1) +
    geom_smooth(se = FALSE, method = "lm", color = "black") + 
    labs(y = "Visits") +
    scale_y_continuous(labels = comma) +
    theme_classic() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))


by_mth_pct_chg <- df %>%
    group_by(month) %>%
    summarize(visits = sum(uk_visits)) %>%
    ungroup() %>%
    arrange(month) %>%
    mutate(pct_change = (visits / lag(visits) - 1) * 100,
           pct_change = ifelse(is.na(pct_change), 0, pct_change)) %>%
    ggplot(aes(x = month, y = pct_change)) +
    geom_point(color = "blue", size = 2) +
    geom_line(color = "blue", size = 1) + 
    geom_smooth(se = FALSE, method = "lm", color = "black") +
    labs(y = "Percent Change",
         x = "Month") +
    theme_classic() + 
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12))

ggarrange(by_mth, 
          by_mth_pct_chg, 
          ncol = 1, 
          align = "hv")

```

